Scalar Items in Embedded Position: {A}n Experimental Revisit

### 2012-07-31 text/paper/paper1.tex
lala

\noindent Given this relative abundance of theoretically conceivable
readings, two interlocked empirical questions arise:
\begin{enumerate}[Q1:]
\item which of these three
conceivable readings are available to na\"{i}ve subjects; and
\item which of the available readings are preferred.\fn{It may
    ultimately be impossible to keep unavailability and strong
    dispreference strictly apart. For our purposes this is not
    important. Every time we speak of availability, the so-inclined
    reader may think of non-negligible preference.}
\end{enumerate}
Addressing these questions empirically is relevant because they lie at
the heart of the current debate about the exact location and nature of
the interface between semantics and pragmatics. The controversy
concerns the question to what extent the upper-bounding inference from
\emph{some} to \emph{some but not all} is conventionalized within the
compositional computation of semantic values. For that matter, it is
particularly relevant to assess the availability and relative
preference of local readings. (Section~\ref{sec:theories-predictions}
will discuss the relevant theoretical positions and the significance
of alleged local readings in more detail.)

@@

A number of empirical studies have already examined the availability
of readings for \as- and \es-sentences
\citep[e.g.][]{GeurtsPouscoulous2009:Embedded-Implic,CliftonDube2010:Embedded-Implic,ChemlaSpector2010:Experimental-Ev}. However,
results have not been as clear-cut as one might have hoped for: for
instance, the available empirical evidence is inconclusive as to
whether local readings exist \citep[see also][for related
discussion]{Tielvan-Tiel2012:Embedded-Scalar} and do not clearly
address all preference relations between readings relevant to
distinguish between theoretical positions. In the following we will
argue that these heterogeneous results are partly due to the specific
tasks used in previous studies, i.e., complexity differences in
variants of the picture-verification paradigm.
% We hypothesized that previous studies might be
% insufficiently informative because of the focus on (variants of) a
% picture-verification paradigm.\dn{relate to Bob van Tiel's work} The
% problem is that in order to test the availability of different
% candidate meanings different pictures have to be presented, so that
% effects of pictorial complexity or stereotypicality could never be
% ruled out entirely.
Moreover, previous studies have only accumulated limited evidence
pertaining to the second question that may help decide between
theoretical positions, namely which of the attested readings subjects
prefer. Finally, previous studies presented target sentences
visually. But as it is often argued that intonational stress on an
embedded scalar item can favor a local reading
\citep[e.g.][]{Horn2006:The-Border-Wars,Geurts2009:Scalar-Implicat,ChemlaSpector2010:Experimental-Ev,Geurts2010:Quantity-Implic,Tielvan-Tiel2012:Embedded-Scalar},
it seems important to present target sentences auditorily, so as to be
able to control for effects of ``silent intonation'' that subjects may
apply when reading a sentence (see Bader 1998, Fodor 2000)\dn{provide
  references}.

In reaction to this situation, we therefore presented sentence
material auditorily and also employed a different kind of visual
presentation of the pictorial material: subjects were presented with
pictures that were initially covered and could incrementally be
uncovered at the subjects' request; at each step of uncovering,
subjects had to decide whether they could already give a truth-value
judgement or needed more information (Comry ????)\dn{insert
  references}. This way we obtained behavioral data that is both
indicative of the reading subjects assumed and independent of the
complexity of the visual stimulus. At the same time, we hypothesized
that the incremental nature of this task would shed light on the
preferences over readings, because the temporal distribution of
truth-value judgments would give away which reading subjects were
waiting to evaluate, so to speak. Adding to previous studies, we
therefore made sure that our method is able to reveal information
about preferential readings by including ambiguous test items like
(\ref{bsp:target-related-filler}) which are known to preferentially
receive the late-closure reading in
(\ref{bsp:target-related-filler-LC}) and not the dispreferred, but
attested early-closure reading (\ref{bsp:target-related-filler-EC})
(Frazier 1987 XYZ).\dn{insert proper references}
\begin{exe}
\ex \label{bsp:target-related-filler} The letter is connected with circles and squares with
  suns.
  \begin{xlist}
  \ex \label{bsp:target-related-filler-LC} The letter is connected
    with squares with suns and circles. \hfill (\LC)
  \ex \label{bsp:target-related-filler-EC} The letter is connected
    with circles with suns and squares with suns. \hfill (\EC)
  \end{xlist}
\end{exe}

@@

\begin{itemize}
\item summarize results
\end{itemize}

The paper is structured as follows. Section~\ref{sec:get-know-your}
elaborates on the three kinds of relevant readings for our target
sentences. Section~\ref{sec:theories-predictions} works out the
different theoretical positions and their predictions about
availability and preference. Section~\ref{sec:previous-studies} recaps
the results of previous studies on this subject, arguing for the need
of a more refined methodology. Section~\ref{sec:design} describes our
experimental design. Section~\ref{sec:results} states our results,
which we discuss in Section~\ref{sec:discussion}.\dn{rephrase
  eventually}

@@

Does traditionalism also predict local readings to be available?  The
answer is slightly different for \as- and the \es-sentences. For
\es-sentences, traditionalism does not predict that a local reading is
available, at least not as a quantity implicature
\citep[c.f.][]{GeurtsPouscoulous2009:Embedded-Implic,ChemlaSpector2010:Experimental-Ev}. This
is because traditionalism assumes that quantity implicatures are
pragmatic enrichments of the literal meaning of an utterance, obtained
by conjoining the literal meaning with a suitable set of negated
alternatives. But since the literal and the local reading of
\es-sentences are logically independent, there is no way that local
readings can be derived in a traditionalist manner as a quantity
implicature: if $X$ and $Y$ are logically independent propositions,
then there is no proposition $Z$ such that $X$ would be equivalent to
$Y \wedge \neg Z$. Traditionalism often does concede that (something
like) local readings can occur if scalar items are marked with
\emph{special intonation}, albeit then as a signal of a different
pragmatic process
\citep[e.g.][]{Horn2006:The-Border-Wars,Geurts2009:Scalar-Implicat,Geurts2010:Quantity-Implic}. We
will come back to this issue in Section~\ref{sec:role-intonation}.

@@

\citet{GeurtsPouscoulous2009:Embedded-Implic} conducted a
picture-verification task to find out whether local readings of \as-
and \es-sentences are available.\fn{Actually,
  \citet{GeurtsPouscoulous2009:Embedded-Implic} did not use
  \es-sentences, but sentences where scalar \emph{some} was embedded
  under non-monotonic quantifier \emph{exactly two} (with appropriate
  pictures, of course). This case is a little more complex, but we
  will gloss over this here, treating their data, as if it was obtained
  for \es-sentences.} The critical conditions of their study presented
subjects with pictures like those in Figure~\ref{fig:weak} and
\ref{fig:local} where the local reading gets a different truth-value
from the literal and the global reading. In particular, for
\as-sentences the local reading is false for the critical picture in
Figure~\ref{fig:weak} whereas the literal and global readings are
true; for \es-sentences the local reading is true for the critical
picture in Figure~\ref{fig:local}, whereas the literal and global
readings are false.\dn{Petra wants more detail in the description of
  this study; I'm not sure that this is needed...?}

@@

\begin{minipage}{6cm}
     \subfloat[][literal]{
    \label{fig:EA-LIT}
        \includegraphics[width=5.5cm]{../pictures/Umweltschuetzer09.eps}
  }

  \subfloat[test][global]{
    \label{fig:EA-LITGLB}
    \includegraphics[width=5.5cm]{../pictures/Umweltschuetzer10.eps}
  }  
   \end{minipage}
   \begin{minipage}{6cm}
     \subfloat[][false1]{
    \label{fig:EA-FALSE1}
        \includegraphics[width=5.5cm]{../pictures/Umweltschuetzer11.eps}
  }

  \subfloat[test][false2]{
    \label{fig:EA-FALSE2}
    \includegraphics[width=5.5cm]{../pictures/Umweltschuetzer12.eps}
  }  
   \end{minipage}
        \caption{}
\label{fig:EA}
\end{figure}

@@

     \subfloat[][\as]{
    \label{fig:AEBars}
        \includegraphics[width=5cm]{../pictures/JudgmentsAllSomeBars.eps}
  }
  \subfloat[test][\es]{
    \label{fig:GEBars}
    \includegraphics[width=5cm]{../pictures/JudgmentsExactlySomeBars.eps}
  }

        \caption{Mean judgments}
\label{fig:BarsPilot}
\label{fig:EA}
\end{figure}

@@

\paragraph{The Target Conditions.} More concretely, our study
presented participants with pictures, like those in
Figures~\ref{fig:InitialPicture}, \ref{fig:exseqAS} or
\ref{fig:exseqES}, depicting a set of four identical central elements
(e.g., letters), which could be connected to surrounding elements
(e.g., triangles). Initially, any potential connections between
central and surrounding elements were covered by dark gray color (see
Figure~\ref{fig:InitialPicture}).
%
\begin{figure}
  \centering
      \fbox{\includegraphics[width=5cm]{../pictures/paper/ae_01_1.pdf}}
      \caption{Initial picture presented to participants where all
        possible connections from each letter to its surrounding
        triangles are covered.}
  \label{fig:InitialPicture}
\end{figure}
%
Sentences to be judged were in German (see (\ref{ex:as}) and
(\ref{ex:es}) below) and presented auditorily, and participants were
asked to listen to a given sentence and then uncover the potential
links step by step until they felt able to give a truth value
judgment. Three options were available for participants at each step:
(i), demand more information, (ii), judge the sentence as true, and
(iii), judge the sentence as false. Importantly, each of the three
potentially available readings corresponded to a specific step in the
uncovering process where the truth value of that reading (and only of
that reading) could be assessed for the first time. The location in
the sequence and the corresponding truth-value judgment differed
between \as- and \es-sentences, as described presently. (This is,
again, because of the logical dependencies between readings.)

@@

\paragraph{Controls.} In order to make sure that subjects understood
the task, i.e., gave truth-value judgements neither sooner not later
than at the first possible position in a sequence, we included a set
of control conditions. These also controlled for response biases that
are independent of the sentence meaning, but only depend on the
experimental procedure or on the picture materials. For each of the
three readings in the \as- and \es-conditions, we constructed an
unambiguous control sentence as in (\ref{bsp:controls-as}) and
(\ref{bsp:controls-es}), requiring the same judgement at the identical
position in the same sequence used for the respective targets. For
instance, example (\ref{bsp:controls-as-1}) requires a
\emph{true}-response analogous to the \as-sentence in (\ref{ex:as})
under its literal reading, (\ref{bsp:controls-as-2}) corresponds to
the \as-sentence under its global reading and
(\ref{bsp:controls-as-3}) corresponds to its local reading. With
regard to the \es-sentences, controls like (\ref{bsp:controls-es-1})
correspond to the global reading, (\ref{bsp:controls-es-2}) to the
literal and (\ref{bsp:controls-es-3}) to the local reading. If,
independently of sentence meaning, there was any bias to respond in a
certain way at any point in the sequences, this should affect control
sentences to the same degree as it affects target conditions.

@@

\paragraph{Controlling ambiguity resolution.} In order to control for
artificial effects of presentation order on the availability of
readings, we included a second type of control conditions. In
particular, we tested (i), whether the order of given readings affects
preferences for ambiguous sentences and (ii), whether prosodic
information can, in principle, shift reading preferences in the
present task. Note that for our target sentences, the logical
entailment relations between readings always require local readings to
be evaluated at the end of a sequence. It could thus in principle be
possible that no subject ever reaches this point, because they gave
truth-value judgements earlier, thereby ending the trial. In that case
it would be unclear if these decisions have been affected by a general
inavailability of local readings or by the fact that one of the
earlier presented readings is the preferred one. By including globally
ambiguous structures with a known preference over available readings,
we thus intended to test whether participants in our task occasionally
choose dispreferred readings, even if they were presented following
preferred readings.

We therefore included globally ambiguous \emph{late-closure
  structures} as in (\ref{bsp:target-related}), which have been
repeatedly shown to exhibit interpretive preferences.
\begin{exe}
\ex \gll Der Brief ist mit Kreisen und Vierecken mit Sonnen
  verbunden. \label{bsp:target-related}\\
The letter is with circles and squares with suns connected.\\
The letter is connected with circles and squares with suns.
\begin{xlist}
  \ex \label{bsp:target-related-LC} The letter is connected with squares containing suns, and it is
    also connected with circles. \hfill{(\lc)}
  \ex \label{bsp:target-related-EC} The letter is connected with circles and squares, both of which
    are containing suns.  \hfill{(\ec)} 
\end{xlist}
\end{exe}
In late-closure sentences, a modifier can be attached to one of two
preceding hosts. More specifically, in (\ref{bsp:target-related}), the
prepositional phrase \emph{with suns} can be attached to the preceding
noun \emph{squares}, resulting in the so-called \emph{late-closure} or
\emph{low-attachment} reading ((\ref{bsp:target-related-LC}), see
Frazier, 1987, for details)\dn{references}. Alternatively, the
\acro{pp} can be attached to the whole conjoined \acro{np}
\emph{circles and squares}, corresponding to the \emph{early-closure}
or \emph{high-attachment} reading
(\ref{bsp:target-related-EC}). Generally, the late-closure reading is
preferred over the early closure reading for sentences comparable to
those in (\ref{bsp:target-related}) (REFS)\dn{references}.

@@

@@

>>>>

\noindent Given this relative abundance of theoretically conceivable
readings, two interlocked empirical questions arise:
\begin{enumerate}[Q1:]
\item which of these three
conceivable readings are available to na\"{i}ve subjects; and
\item which of the available readings are preferred.\fn{It may
    ultimately be impossible to keep unavailability and strong
    dispreference strictly apart. For our purposes this is not
    important. Every time we speak of availability, the so-inclined
    reader may think of non-negligible preference.}
\end{enumerate}
Addressing these questions empirically is relevant because they lie at
the heart of the current debate about the exact location and nature of
the interface between semantics and pragmatics. The controversy
concerns the question to what extent the upper-bounding inference from
\emph{some} to \emph{some but not all} is conventionalized within the
compositional computation of semantic values. For that matter, it is
particularly relevant to assess the availability and relative
preference of local readings. %(Section~\ref{sec:theories-predictions}
%will discuss the relevant theoretical positions and the significance
%of alleged local readings in more detail.)

@@

A number of empirical studies have already examined the availability
of readings for \as- and \es-sentences
\citep[e.g.][]{GeurtsPouscoulous2009:Embedded-Implic,CliftonDube2010:Embedded-Implic,ChemlaSpector2010:Experimental-Ev}. However,
results have not been as clear-cut as one might have hoped for: for
instance, the available empirical evidence is inconclusive as to
whether local readings exist \citep[see also][for related
discussion]{Tielvan-Tiel2012:Embedded-Scalar} and do not clearly
address all preference relations between readings relevant to
distinguish between theoretical positions. In the following, we will
argue that these heterogeneous results were partly due to the use of 
standard procedures such as picture-verification paradigm, which might not 
be sufficiently sensitive for the phenomenon under investigation,
for instance, due to possible confounding effects of the pictures used.

% We hypothesized that previous studies might be
% insufficiently informative because of the focus on (variants of) a
% picture-verification paradigm.\dn{relate to Bob van Tiel's work} The
% problem is that in order to test the availability of different
% candidate meanings different pictures have to be presented, so that
% effects of pictorial complexity or stereotypicality could never be
% ruled out entirely.
Moreover, previous studies have only accumulated limited evidence
pertaining to the second question that may help decide between
theoretical positions, namely which of the attested readings subjects
prefer. Finally, previous studies presented target sentences
visually. But as it is often argued that intonational stress on an
embedded scalar item can favor a local reading
\citep[e.g.][]{Horn2006:The-Border-Wars,Geurts2009:Scalar-Implicat,ChemlaSpector2010:Experimental-Ev,Geurts2010:Quantity-Implic,Tielvan-Tiel2012:Embedded-Scalar},
it seems important to present target sentences auditorily, so as to be
able to control for effects of ``silent intonation'' that subjects may
apply when reading a sentence (see Bader 1998, Fodor 2000)\dn{provide
  references}.

In reaction to this situation, we therefore presented sentence
material auditorily and also employed a different kind of visual
presentation of the pictorial material.
%: subjects were presented with
%pictures that were initially covered and could incrementally be
%uncovered at the subjects' request; at each step of uncovering,
%subjects had to decide whether they could already give a truth-value
%judgement or needed more information (Comry ????)\dn{insert
%  references}. 
This way we obtained behavioral data that is both
indicative of the reading subjects assumed and independent of the
confounding pictorial effects. 
%At the same time, we hypothesized
%that the incremental nature of this task would shed light on the
%preferences over readings, because the temporal distribution of
%truth-value judgments would give away which reading subjects were
%waiting to evaluate, so to speak. 
Adding to previous studies, we made sure that our method is able 
to reveal information about reading preferences by relating our 
targets to strucutres that are known in the psycholinguistic 
literature for exhibiting certain preferences.

%including ambiguous ambiguous structures, 
%which have been shown to exhibit certain preferences in previous

% like
%(\ref{bsp:target-related-filler}) which are known to preferentially
%receive the late-closure reading in
%(\ref{bsp:target-related-filler-LC}) and not the dispreferred, but
%attested early-closure reading (\ref{bsp:target-related-filler-EC})
%(Frazier 1987 XYZ).\dn{insert proper references}
%\begin{exe}
%\ex \label{bsp:target-related-filler} The letter is connected with circles and squares with
%  suns.
%  \begin{xlist}
%  \ex \label{bsp:target-related-filler-LC} The letter is connected
%    with squares with suns and circles. \hfill (\LC)
%  \ex \label{bsp:target-related-filler-EC} The letter is connected
%    with circles with suns and squares with suns. \hfill (\EC)
%  \end{xlist}
%\end{exe}

@@

%\begin{itemize}
%\item summarize results
%\end{itemize}

%The paper is structured as follows. Section~\ref{sec:get-know-your}
%elaborates on the three kinds of relevant readings for our target
%sentences. Section~\ref{sec:theories-predictions} works out the
%different theoretical positions and their predictions about
%availability and preference. Section~\ref{sec:previous-studies} recaps
%the results of previous studies on this subject, arguing for the need
%of a more refined methodology. Section~\ref{sec:design} describes our
%experimental design. Section~\ref{sec:results} states our results,
%which we discuss in Section~\ref{sec:discussion}.\dn{rephrase
%  eventually}

@@

Does traditionalism also predict local readings to be available?  The
answer is slightly different for \as- and the \es-sentences. For
\es-sentences, traditionalism does not predict that a local reading is
available, at least not as a quantity implicature
\citep[c.f.][]{GeurtsPouscoulous2009:Embedded-Implic,ChemlaSpector2010:Experimental-Ev}. This
is because traditionalism assumes that quantity implicatures are
pragmatic enrichments of the literal meaning of an utterance, obtained
by conjoining the literal meaning with a suitable set of negated
alternatives. But since the literal and the local reading of
\es-sentences are logically independent, there is no way that local
readings can be derived in a traditionalist manner as a quantity
implicature: if $X$ and $Y$ are logically independent propositions,
then there is no proposition $Z$ such that $X$ would be equivalent to
$Y \wedge \neg Z$. Traditionalism often does concede that (something
like) local readings can occur if scalar items are marked with
\emph{special accentuation}, albeit then as a signal of a different
pragmatic process
\citep[e.g.][]{Horn2006:The-Border-Wars,Geurts2009:Scalar-Implicat,Geurts2010:Quantity-Implic}. We
will come back to this issue in Section~\ref{sec:role-intonation}.

@@

\citet{GeurtsPouscoulous2009:Embedded-Implic} conducted a
picture-verification task to find out whether local readings of \as-
and \es-sentences are available.\fn{Actually,
  \citet{GeurtsPouscoulous2009:Embedded-Implic} did not use
  \es-sentences, but sentences where scalar \emph{some} was embedded
  under non-monotonic quantifier \emph{exactly two} (with appropriate
  pictures, of course). This case is a little more complex, but we
  will gloss over this here, treating their data, as if it was obtained
  for \es-sentences.} Subjects were presented with pictures like 
  those in Figure~\ref{fig:weak} and
\ref{fig:local} where the local reading gets a different truth-value
from the literal and the global reading. For
\as-sentences, the local reading is false for the critical picture in
Figure~\ref{fig:weak} whereas the literal and global readings are
true; for \es-sentences the local reading is true for the critical
picture in Figure~\ref{fig:local}, whereas the literal and global
readings are false.\dn{Petra wants more detail in the description of
  this study; I'm not sure that this is needed...?}

@@

%\begin{figure}[t]
%  \centering
% 
%\begin{minipage}{6cm}
%     \subfloat[][literal]{
%    \label{fig:EA-LIT}
%        \includegraphics[width=5.5cm]{../pictures/Umweltschuetzer09.eps}
%  }
%
%  \subfloat[test][global]{
%    \label{fig:EA-LITGLB}
%    \includegraphics[width=5.5cm]{../pictures/Umweltschuetzer10.eps}
%  }  
%   \end{minipage}
%   \begin{minipage}{6cm}
%     \subfloat[][false1]{
%    \label{fig:EA-FALSE1}
%        \includegraphics[width=5.5cm]{../pictures/Umweltschuetzer11.eps}
%  }
%
%  \subfloat[test][false2]{
%    \label{fig:EA-FALSE2}
%    \includegraphics[width=5.5cm]{../pictures/Umweltschuetzer12.eps}
%  }  
%   \end{minipage}
%        \caption{}
%\label{fig:EA}
%\end{figure}

@@

%\begin{figure}[t]
%  \centering
% 
%
%     \subfloat[][\as]{
%    \label{fig:AEBars}
%        \includegraphics[width=5cm]{../pictures/JudgmentsAllSomeBars.eps}
%  }
%  \subfloat[test][\es]{
%    \label{fig:GEBars}
%    \includegraphics[width=5cm]{../pictures/JudgmentsExactlySomeBars.eps}
%  }  
%  
%        \caption{Mean judgments}
%\label{fig:BarsPilot}
%\label{fig:EA}
%\end{figure}

@@

\paragraph{Target Conditions.} More concretely, our study
presented participants with pictures, like those in
Figures~\ref{fig:InitialPicture}, \ref{fig:exseqAS} or
\ref{fig:exseqES}, depicting a set of four identical central elements
(e.g., letters), which could be connected to surrounding elements
(e.g., triangles). Initially, any potential connections between
central and surrounding elements were covered by dark gray color (see
Figure~\ref{fig:InitialPicture}).
%
\begin{figure}
  \centering
      \fbox{\includegraphics[width=5cm]{../pictures/paper/ae_01_1.pdf}}
      \caption{Initial picture presented to participants where all
        possible connections from each letter to its surrounding
        triangles are covered.}
  \label{fig:InitialPicture}
\end{figure}
%
Sentences to be judged were in German (see (\ref{ex:as}) and
(\ref{ex:es}) below) and presented auditorily, and participants were
asked to listen to a given sentence and then uncover the potential
links step by step until they felt able to give a truth value
judgment. Three options were available for participants at each step:
(i), demand more information, (ii), judge the sentence as true, and
(iii), judge the sentence as false. Importantly, each of the three
potentially available readings corresponded to a specific step in the
uncovering process where the truth value of that reading (and only of
that reading) could be assessed for the first time. The location in
the sequence and the corresponding truth-value judgment differed
between \as- and \es-sentences, as described presently. (This is,
again, because of the logical dependencies between readings.)

@@

\paragraph{Positional controls.} In order to make sure that subjects understood
the task, i.e., gave truth-value judgements neither sooner not later
than at the first possible position in a sequence, we included a set
of control conditions. These also controlled for response biases that
are independent of the sentence meaning, but only depend on the
experimental procedure or on the picture materials. For each of the
three readings in the \as- and \es-conditions, we constructed an
unambiguous control sentence as in (\ref{bsp:controls-as}) and
(\ref{bsp:controls-es}), requiring the same judgement at the identical
position in the same sequence used for the respective targets. For
instance, example (\ref{bsp:controls-as-1}) requires a
\emph{true}-response analogous to the \as-sentence in (\ref{ex:as})
under its literal reading, (\ref{bsp:controls-as-2}) corresponds to
the \as-sentence under its global reading and
(\ref{bsp:controls-as-3}) corresponds to its local reading. With
regard to the \es-sentences, controls like (\ref{bsp:controls-es-1})
correspond to the global reading, (\ref{bsp:controls-es-2}) to the
literal and (\ref{bsp:controls-es-3}) to the local reading. If,
independently of sentence meaning, there was any bias to respond in a
certain way at any point in the sequences, this should affect control
sentences to the same degree as it affects target conditions.

@@

\paragraph{Preference-related controls} In order to control for
artificial effects of presentation order on the availability of
readings, we included a second type of control conditions. In
particular, we tested (i), whether the order of given readings affects
preferences for ambiguous sentences and (ii), whether prosodic
information can, in principle, shift reading preferences in the
present task. Note that for our target sentences, the logical
entailment relations between readings always require local readings to
be evaluated at the end of a sequence. It could thus in principle be
possible that no subject ever reaches this point, because they gave
truth-value judgements earlier, thereby ending the trial. In that case
it would be unclear if these decisions have been affected by a general
inavailability of local readings or by the fact that one of the
earlier presented readings is the preferred one. By including globally
ambiguous structures with a known preference over available readings,
we thus intended to test whether participants in our task occasionally
choose dispreferred readings, even if they were presented subsequent 
to preferred readings.

We therefore included globally ambiguous \emph{late-closure
  structures} as in (\ref{bsp:target-related}), which have been
repeatedly shown to exhibit interpretive preferences.
\begin{exe}
\ex \gll Der Brief ist mit Kreisen und Vierecken mit Sonnen
  verbunden. \label{bsp:target-related}\\
The letter is with circles and squares with suns connected.\\
The letter is connected with circles and squares with suns.
\begin{xlist}
  \ex \label{bsp:target-related-LC} The letter is connected with squares containing suns, and it is
    also connected with circles. \hfill{(\lc)}
  \ex \label{bsp:target-related-EC} The letter is connected with circles and squares, both of which
    are containing suns.  \hfill{(\ec)} 
\end{xlist}

\ex \gll Der Brief ist mit Kreisen, die Sonnen beinhalten und Vierecken
  verbunden. \label{bsp:target-related2}\\
The letter is with circles, which suns contain and squares connected.\\
The letter is connected with circles containing suns, and with squares.

\end{exe}
In late-closure sentences, a modifier can be attached to one of two
preceding hosts. More specifically, in (\ref{bsp:target-related}), the
prepositional phrase \emph{with suns} can be attached to the preceding
noun \emph{squares}, resulting in the so-called \emph{late-closure} or
\emph{low-attachment} reading ((\ref{bsp:target-related-LC}), see
Frazier, 1987, for details)\dn{references}. Alternatively, the
\acro{pp} can be attached to the whole conjoined \acro{np}
\emph{circles and squares}, corresponding to the \emph{early-closure}
or \emph{high-attachment} reading
(\ref{bsp:target-related-EC}). Generally, the late-closure reading is
preferred over the early closure reading for sentences comparable to
those in (\ref{bsp:target-related}) (REFS)\dn{references}.

@@

\paragraph{\as- and \es-sentences and positional controls}

We constructed a set of 15 items for \as- and \es-sentences, respectively,
analogous to the examples in ... and ..

Sentences were introduced by a subject DP, including the quantifiers 
{\it alle diese} ("all of these") or {\it genau einer der} ("exactly 
one of these"), as well as the head noun which denoted different icons, 
such as letters or bells. Subjects were followed by an auxiliary, and 
a PP containing {\it einige}, a possessive pronoun, and a noun denoting 
a geometrical object, e.g. a triangle or a square. Each sentence was recorded 
with a stressed and an unstressed version of scalar "some" (see section xx 
below). As described in section ..., three control sentences were created 
for each experimental item, corresponding to the critical positions.

Items were evenly distributed across 5 lists using a Latin square design.

\paragraph{Preference-related controls}
For controlling preferences, 28 ambiguous Late closure sentences and 28 
of their disambiguated counterparts were constructed as described in Section ... 
In these sentences, subject DPs were always denoting icons and were followed
by an auxiliary. In the ambiguous sentences, the auxiliary was followed by two PPs.
The first of these PPs consisted in a coordination of two geometrical 
objects, whereas the second PP denoted an icon again. In contrast to these 
sentences, the first part of the conjunct in the unambiguous sentences,was 
modified by a relative clause.

\paragraph{Acoustic properties}

@@

In sum, our speaker reliably produced (i) differences in accent realization for the target sentences and (ii) different boundary realizations for the target- related fillers.
Whereas accented elements clearly differed from their unac-
cented counterpart by showing an increase in duration and F0 range, prosodic
boundaries for our target-related fillers were realized by pre-final lengthening
(i.e. an increase in F0 and duration at the position preceding the boundary).
